---
title: "Step 1: Internal Reliability Analysis"
output:
  html_notebook:
    toc: true
    toc_float: true
---

This analysis evaluates internal reliability of each construct and identifies
items for potential removal based on psychometric criteria only.

**Criteria for item removal:**

- Item-total correlation < 0.3

**We do NOT remove items based on:**

- Whether they produce significant path coefficients
- Whether the model converges

```{r setup, message=FALSE, warning=FALSE}
library(ltm)
library(dplyr)
library(ggplot2)
library(reshape2)
library(knitr)
library(kableExtra)

theme_set(theme_minimal(base_size = 12))
```

# Load Data

```{r}
df <- read.csv("../data/Dec2BonusDataset.csv")

# Reverse-code negatively worded items
reverse_items <- c("Q2", "Q7", "Q9", "Q15", "Q25", "Q31", "Q37", "Q42")
for (item in reverse_items) {
  df[[item]] <- 6 - df[[item]]
}

df_items <- df[paste0("Q", 0:44)]

cat("Sample size:", nrow(df_items), "\n")
cat("Number of items:", ncol(df_items), "\n")
```

# Define Constructs

```{r}
constructs <- list(
  system_quality = paste0("Q", 0:5),
  service_quality = paste0("Q", 6:15),
  perceived_fairness = paste0("Q", 16:23),
  perceived_ease_of_use = paste0("Q", 24:28),
  perceived_usefulness = paste0("Q", 29:35),
  attitude_toward_use = paste0("Q", 36:39),
  behavioral_intention = paste0("Q", 40:44)
)

construct_labels <- c(
  system_quality = "System Quality",
  service_quality = "Service Quality",
  perceived_fairness = "Perceived Fairness",
  perceived_ease_of_use = "Perceived Ease of Use",
  perceived_usefulness = "Perceived Usefulness",
  attitude_toward_use = "Attitude Toward Use",
  behavioral_intention = "Behavioral Intention"
)
```

# Reliability Analysis

```{r}
analyze_reliability <- function(data, items, construct_name) {
  subset <- data[, items]

  # Cronbach's alpha
  alpha <- cronbach.alpha(subset)$alpha

  # Item-total correlations (corrected)
  total_score <- rowSums(subset)
  item_total_cors <- sapply(items, function(item) {
    cor(subset[[item]], total_score - subset[[item]])
  })

  # Alpha if item deleted
  alpha_if_deleted <- sapply(items, function(item) {
    remaining <- setdiff(items, item)
    if (length(remaining) >= 2) {
      cronbach.alpha(data[, remaining])$alpha
    } else {
      NA
    }
  })

  result <- data.frame(
    item = items,
    item_total_cor = item_total_cors,
    alpha_if_deleted = alpha_if_deleted,
    flag = item_total_cors < 0.3,
    stringsAsFactors = FALSE
  )

  return(list(alpha = alpha, details = result))
}

# Run analysis for all constructs
reliability_results <- list()
for (name in names(constructs)) {
  reliability_results[[name]] <- analyze_reliability(
    df_items,
    constructs[[name]],
    name
  )
}
```

## Summary Table

```{r}
summary_df <- data.frame(
  Construct = construct_labels[names(constructs)],
  Items = sapply(constructs, length),
  Alpha = sapply(reliability_results, function(x) x$alpha),
  Flagged = sapply(reliability_results, function(x) sum(x$details$flag))
)

summary_df %>%
  mutate(
    Alpha = round(Alpha, 3),
    Status = case_when(
      Alpha >= 0.7 ~ "Good",
      Alpha >= 0.6 ~ "Acceptable",
      TRUE ~ "Poor"
    )
  ) %>%
  kable(align = "lcccc") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## Cronbach's Alpha Visualization

```{r fig.width=8, fig.height=5}
alpha_df <- data.frame(
  construct = factor(construct_labels[names(constructs)],
                     levels = rev(construct_labels[names(constructs)])),
  alpha = sapply(reliability_results, function(x) x$alpha)
)

ggplot(alpha_df, aes(x = construct, y = alpha)) +
  geom_col(aes(fill = alpha >= 0.6), width = 0.7) +
  geom_hline(yintercept = 0.6, linetype = "dashed", color = "red", size = 1) +
  geom_hline(yintercept = 0.7, linetype = "dashed", color = "darkgreen", size = 1) +
  geom_text(aes(label = round(alpha, 2)), hjust = -0.2, size = 4) +
  coord_flip() +
  scale_fill_manual(values = c("TRUE" = "steelblue", "FALSE" = "coral"),
                    guide = "none") +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
  labs(
    title = "Cronbach's Alpha by Construct",
    subtitle = "Red line = 0.6 (acceptable), Green line = 0.7 (good)",
    x = NULL,
    y = "Cronbach's Alpha"
  ) +
  theme(plot.title = element_text(face = "bold"))
```

## Item-Total Correlations

```{r fig.width=10, fig.height=8}
# Combine all item-total correlations
all_items <- do.call(rbind, lapply(names(reliability_results), function(name) {
  df <- reliability_results[[name]]$details
  df$construct <- construct_labels[name]
  df
}))

ggplot(all_items, aes(x = reorder(item, item_total_cor), y = item_total_cor)) +
  geom_col(aes(fill = flag), width = 0.8) +
  geom_hline(yintercept = 0.3, linetype = "dashed", color = "red", size = 0.8) +
  facet_wrap(~construct, scales = "free_y", ncol = 2) +
  coord_flip() +
  scale_fill_manual(
    values = c("TRUE" = "coral", "FALSE" = "steelblue"),
    labels = c("TRUE" = "Below threshold", "FALSE" = "Acceptable"),
    name = "Status"
  ) +
  labs(
    title = "Item-Total Correlations by Construct",
    subtitle = "Red line = 0.3 threshold; items below are flagged for removal",
    x = NULL,
    y = "Corrected Item-Total Correlation"
  ) +
  theme(
    plot.title = element_text(face = "bold"),
    legend.position = "bottom"
  )
```

## Detailed Results by Construct

```{r results='asis'}
for (name in names(reliability_results)) {
  cat("\n\n### ", construct_labels[name], "\n\n")

  res <- reliability_results[[name]]
  cat("**Cronbach's Alpha:**", round(res$alpha, 3), "\n\n")

  res$details %>%
    mutate(
      item_total_cor = round(item_total_cor, 3),
      alpha_if_deleted = round(alpha_if_deleted, 3),
      flag = ifelse(flag, "REMOVE", "Keep")
    ) %>%
    rename(
      Item = item,
      `Item-Total r` = item_total_cor,
      `Alpha if Deleted` = alpha_if_deleted,
      Decision = flag
    ) %>%
    kable() %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
    print()
}
```

# Items to Remove

Based strictly on item-total correlation < 0.3:

```{r}
items_to_remove <- all_items$item[all_items$flag]

if (length(items_to_remove) > 0) {
  cat("Items flagged for removal:\n")
  for (item in items_to_remove) {
    construct <- all_items$construct[all_items$item == item]
    cor_val <- round(all_items$item_total_cor[all_items$item == item], 3)
    cat("  -", item, "(", construct, "), r =", cor_val, "\n")
  }
} else {
  cat("No items flagged for removal.\n")
}
```

# Revised Constructs

```{r}
constructs_revised <- lapply(names(constructs), function(name) {
  items <- constructs[[name]]
  flagged <- reliability_results[[name]]$details
  keep <- flagged$item[!flagged$flag]
  as.character(keep)
})
names(constructs_revised) <- names(constructs)

# Summary of changes
cat("\nConstruct changes:\n\n")
for (name in names(constructs)) {
  original <- length(constructs[[name]])
  revised <- length(constructs_revised[[name]])
  removed <- original - revised
  cat(construct_labels[name], ":", original, "->", revised, "items")
  if (removed > 0) {
    cat(" (removed", removed, ")")
  }
  cat("\n")
}
```

## Revised Alpha Values

```{r}
revised_summary <- data.frame(
  Construct = construct_labels[names(constructs)],
  Original_Items = sapply(constructs, length),
  Original_Alpha = sapply(reliability_results, function(x) x$alpha),
  Revised_Items = sapply(constructs_revised, length),
  Revised_Alpha = sapply(names(constructs_revised), function(name) {
    items <- constructs_revised[[name]]
    if (length(items) >= 2) {
      cronbach.alpha(df_items[, items])$alpha
    } else {
      NA
    }
  })
)

revised_summary %>%
  mutate(
    Original_Alpha = round(Original_Alpha, 3),
    Revised_Alpha = round(Revised_Alpha, 3),
    Change = round(Revised_Alpha - Original_Alpha, 3)
  ) %>%
  kable(
    col.names = c("Construct", "Original N", "Original Alpha",
                  "Revised N", "Revised Alpha", "Change")
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

# Save for Next Step

```{r}
saveRDS(constructs_revised, "constructs_revised.rds")
saveRDS(df_items, "df_items.rds")
cat("\nData saved for CFA analysis.\n")
```
