---
title: "Step 2: Confirmatory Factor Analysis"
output:
  html_notebook:
    toc: true
    toc_float: true
---

This analysis fits the measurement model (CFA) and evaluates:

1. Factor loadings (threshold: >= 0.5, or >= 0.4 with justification)
2. Model fit indices
3. Discriminant validity

Items with factor loadings < 0.4 are removed, then the model is refit.

```{r setup, message=FALSE, warning=FALSE}
library(lavaan)
library(semTools)
library(dplyr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(corrplot)

theme_set(theme_minimal(base_size = 12))
```

# Load Data

```{r}
constructs <- readRDS("constructs_revised.rds")
df_items <- readRDS("df_items.rds")

construct_labels <- c(
  system_quality = "System Quality",
  service_quality = "Service Quality",
  perceived_fairness = "Perceived Fairness",
  perceived_ease_of_use = "Perceived Ease of Use",
  perceived_usefulness = "Perceived Usefulness",
  attitude_toward_use = "Attitude Toward Use",
  behavioral_intention = "Behavioral Intention"
)

cat("Sample size:", nrow(df_items), "\n")
```

# Build CFA Model

```{r}
build_cfa_model <- function(constructs) {
  model_lines <- sapply(names(constructs), function(name) {
    items <- constructs[[name]]
    paste0(name, " =~ ", paste(items, collapse = " + "))
  })
  paste(model_lines, collapse = "\n")
}

cfa_model <- build_cfa_model(constructs)
cat("CFA Model Specification:\n\n")
cat(cfa_model)
```

# Fit Initial CFA

```{r}
fit_cfa <- cfa(
  cfa_model,
  data = df_items,
  std.lv = TRUE,
  estimator = "MLM"  # Robust ML for non-normal data
)

summary(fit_cfa, fit.measures = TRUE, standardized = TRUE)
```

# Model Fit Indices

```{r}
fit_measures <- fitMeasures(fit_cfa, c(
  "chisq", "df", "pvalue",
  "cfi", "tli", "rmsea", "rmsea.ci.lower", "rmsea.ci.upper", "srmr"
))

fit_df <- data.frame(
  Index = c("Chi-square", "df", "p-value", "CFI", "TLI", "RMSEA", "RMSEA 90% CI", "SRMR"),
  Value = c(
    round(fit_measures["chisq"], 2),
    fit_measures["df"],
    round(fit_measures["pvalue"], 4),
    round(fit_measures["cfi"], 3),
    round(fit_measures["tli"], 3),
    round(fit_measures["rmsea"], 3),
    paste0("[", round(fit_measures["rmsea.ci.lower"], 3), ", ",
           round(fit_measures["rmsea.ci.upper"], 3), "]"),
    round(fit_measures["srmr"], 3)
  ),
  Threshold = c("-", "-", "> 0.05", ">= 0.90", ">= 0.90", "<= 0.08", "-", "<= 0.08"),
  Status = c(
    "-", "-",
    ifelse(fit_measures["pvalue"] > 0.05, "Good", "Poor"),
    ifelse(fit_measures["cfi"] >= 0.90, "Good", ifelse(fit_measures["cfi"] >= 0.85, "Acceptable", "Poor")),
    ifelse(fit_measures["tli"] >= 0.90, "Good", ifelse(fit_measures["tli"] >= 0.85, "Acceptable", "Poor")),
    ifelse(fit_measures["rmsea"] <= 0.08, "Good", ifelse(fit_measures["rmsea"] <= 0.10, "Acceptable", "Poor")),
    "-",
    ifelse(fit_measures["srmr"] <= 0.08, "Good", "Poor")
  )
)

fit_df %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

# Factor Loadings

```{r}
loadings <- standardizedSolution(fit_cfa) %>%
  filter(op == "=~") %>%
  select(construct = lhs, item = rhs, loading = est.std, se, pvalue) %>%
  mutate(
    flag = abs(loading) < 0.4,
    construct_label = construct_labels[construct]
  )

loadings %>%
  mutate(
    loading = round(loading, 3),
    se = round(se, 3),
    pvalue = round(pvalue, 4),
    status = ifelse(flag, "REMOVE", "Keep")
  ) %>%
  select(Construct = construct_label, Item = item, Loading = loading,
         SE = se, `p-value` = pvalue, Status = status) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  row_spec(which(loadings$flag), background = "#ffcccc")
```

## Factor Loadings Visualization

```{r fig.width=10, fig.height=8}
ggplot(loadings, aes(x = reorder(item, loading), y = loading)) +
  geom_col(aes(fill = flag), width = 0.8) +
  geom_hline(yintercept = 0.4, linetype = "dashed", color = "red", size = 0.8) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "orange", size = 0.8) +
  facet_wrap(~construct_label, scales = "free_y", ncol = 2) +
  coord_flip() +
  scale_fill_manual(
    values = c("TRUE" = "coral", "FALSE" = "steelblue"),
    labels = c("TRUE" = "Below 0.4", "FALSE" = "Acceptable"),
    name = "Status"
  ) +
  labs(
    title = "Standardized Factor Loadings",
    subtitle = "Red line = 0.4 (minimum), Orange line = 0.5 (preferred)",
    x = NULL,
    y = "Standardized Loading"
  ) +
  theme(
    plot.title = element_text(face = "bold"),
    legend.position = "bottom"
  )
```

# Items to Remove (Loading < 0.4)

```{r}
items_low_loading <- loadings$item[loadings$flag]

if (length(items_low_loading) > 0) {
  cat("Items with factor loading < 0.4:\n\n")
  for (item in items_low_loading) {
    row <- loadings[loadings$item == item, ]
    cat("  -", item, "(", row$construct_label, "), loading =", round(row$loading, 3), "\n")
  }
} else {
  cat("No items have factor loading < 0.4.\n")
}
```

# Discriminant Validity

Inter-construct correlations should be < 0.85 for discriminant validity.

```{r fig.width=8, fig.height=7}
cor_matrix <- lavInspect(fit_cfa, "cor.lv")

# Rename for display
rownames(cor_matrix) <- construct_labels[rownames(cor_matrix)]
colnames(cor_matrix) <- construct_labels[colnames(cor_matrix)]

corrplot(
  cor_matrix,
  method = "color",
  type = "lower",
  addCoef.col = "black",
  tl.col = "black",
  tl.srt = 45,
  col = colorRampPalette(c("steelblue", "white", "firebrick"))(100),
  title = "Inter-Construct Correlations",
  mar = c(0, 0, 2, 0)
)
```

```{r}
# Check for problematic correlations
cor_long <- reshape2::melt(cor_matrix)
cor_long <- cor_long[cor_long$Var1 != cor_long$Var2, ]
high_cors <- cor_long[abs(cor_long$value) > 0.85, ]

if (nrow(high_cors) > 0) {
  cat("WARNING: High correlations (> 0.85) detected:\n\n")
  print(high_cors)
} else {
  cat("All inter-construct correlations < 0.85. Discriminant validity supported.\n")
}
```

# Convergent Validity (AVE)

Average Variance Extracted should be >= 0.5 for convergent validity.

```{r}
# Calculate AVE manually
ave_values <- sapply(names(constructs), function(name) {
  construct_loadings <- loadings$loading[loadings$construct == name]
  mean(construct_loadings^2)
})

ave_df <- data.frame(
  Construct = construct_labels[names(ave_values)],
  AVE = round(ave_values, 3),
  Status = ifelse(ave_values >= 0.5, "Good", "Below threshold")
)

ave_df %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

# Composite Reliability

```{r}
cr_values <- compRelSEM(fit_cfa)

cr_df <- data.frame(
  Construct = construct_labels[names(cr_values)],
  CR = round(as.numeric(cr_values), 3),
  Status = ifelse(as.numeric(cr_values) >= 0.7, "Good",
                  ifelse(as.numeric(cr_values) >= 0.6, "Acceptable", "Poor"))
)

cr_df %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

# Refit After Removing Low-Loading Items

```{r}
if (length(items_low_loading) > 0) {
  cat("Removing items with loading < 0.4:", paste(items_low_loading, collapse = ", "), "\n\n")

  constructs_final <- lapply(names(constructs), function(name) {
    items <- constructs[[name]]
    setdiff(items, items_low_loading)
  })
  names(constructs_final) <- names(constructs)

  # Check if any construct has < 2 items
  for (name in names(constructs_final)) {
    if (length(constructs_final[[name]]) < 2) {
      cat("WARNING:", construct_labels[name], "has fewer than 2 items remaining.\n")
    }
  }

  # Refit model
  cfa_model_final <- build_cfa_model(constructs_final)
  cat("\nRevised CFA Model:\n\n")
  cat(cfa_model_final)

  fit_cfa_final <- cfa(
    cfa_model_final,
    data = df_items,
    std.lv = TRUE,
    estimator = "MLM"
  )

  cat("\n\nRevised Model Fit:\n")
  fit_final <- fitMeasures(fit_cfa_final, c("cfi", "tli", "rmsea", "srmr"))
  print(round(fit_final, 3))

} else {
  constructs_final <- constructs
  fit_cfa_final <- fit_cfa
  cat("No items removed. Using original model.\n")
}
```

# Save for SEM Analysis

```{r}
saveRDS(constructs_final, "constructs_final.rds")
saveRDS(fit_cfa_final, "fit_cfa_final.rds")
cat("\nFinal constructs saved for SEM analysis.\n")
```

# Summary

```{r}
cat("\nFinal construct composition:\n\n")
for (name in names(constructs_final)) {
  cat(construct_labels[name], ":", paste(constructs_final[[name]], collapse = ", "), "\n")
}
```
