---
title: "Step 3: Structural Equation Modeling"
output:
  html_notebook:
    toc: true
    toc_float: true
---

This analysis fits the structural model (SEM) to test the hypothesized causal
paths in the FAIR model. Results are reported regardless of statistical
significance.

```{r setup, message=FALSE, warning=FALSE}
library(lavaan)
library(semPlot)
library(semTools)
library(dplyr)
library(ggplot2)
library(knitr)
library(kableExtra)

theme_set(theme_minimal(base_size = 12))
```

# Load Data

```{r}
constructs <- readRDS("constructs_final.rds")
df_items <- readRDS("df_items.rds")

construct_labels <- c(
  system_quality = "System Quality",
  service_quality = "Service Quality",
  perceived_fairness = "Perceived Fairness",
  perceived_ease_of_use = "Perceived Ease of Use",
  perceived_usefulness = "Perceived Usefulness",
  attitude_toward_use = "Attitude Toward Use",
  behavioral_intention = "Behavioral Intention"
)

cat("Sample size:", nrow(df_items), "\n")
cat("\nFinal constructs:\n")
for (name in names(constructs)) {
  cat("  ", construct_labels[name], ":", length(constructs[[name]]), "items\n")
}
```

# Hypotheses

Based on the FAIR model extending TAM:

| # | Hypothesis | Path |
|---|------------|------|
| H1 | System quality -> Perceived fairness | sy.q -> p.f |
| H2 | Service quality -> Perceived fairness | se.q -> p.f |
| H3 | Perceived fairness -> Perceived usefulness | p.f -> p.u |
| H4 | Perceived fairness -> Attitude | p.f -> a.t.u |
| H5 | System quality -> Perceived ease of use | sy.q -> p.e.o.u |
| H6 | Service quality -> Perceived ease of use | se.q -> p.e.o.u |
| H7 | System quality -> Perceived usefulness | sy.q -> p.u |
| H8 | Service quality -> Perceived usefulness | se.q -> p.u |
| H9 | Perceived ease of use -> Perceived usefulness | p.e.o.u -> p.u |
| H10 | Perceived ease of use -> Attitude | p.e.o.u -> a.t.u |
| H11 | Perceived usefulness -> Attitude | p.u -> a.t.u |
| H12 | Attitude -> Behavioral intention | a.t.u -> b.i |

# FAIR Model (7 Constructs)

## Model Specification

```{r}
build_sem_model <- function(constructs, structural_paths) {
  # Measurement model
  measurement <- sapply(names(constructs), function(name) {
    items <- constructs[[name]]
    paste0(name, " =~ ", paste(items, collapse = " + "))
  })

  paste(c(measurement, "", "# Structural paths", structural_paths), collapse = "\n")
}

fair_paths <- c(
  "# Fairness paths (novel hypotheses)",
  "perceived_fairness ~ system_quality + service_quality",
  "",
  "# TAM paths",
  "perceived_ease_of_use ~ system_quality + service_quality",
  "perceived_usefulness ~ system_quality + service_quality + perceived_fairness + perceived_ease_of_use",
  "attitude_toward_use ~ perceived_fairness + perceived_ease_of_use + perceived_usefulness",
  "behavioral_intention ~ attitude_toward_use"
)

fair_model <- build_sem_model(constructs, fair_paths)
cat("FAIR Model Specification:\n\n")
cat(fair_model)
```

## Fit Model

```{r}
fit_fair <- sem(
  fair_model,
  data = df_items,
  std.lv = TRUE,
  estimator = "MLM"
)
```

## Model Fit

```{r}
fit_measures <- fitMeasures(fit_fair, c(
  "chisq", "df", "pvalue",
  "cfi", "tli", "rmsea", "rmsea.ci.lower", "rmsea.ci.upper", "srmr"
))

fit_df <- data.frame(
  Index = c("Chi-square", "df", "p-value", "CFI", "TLI", "RMSEA", "RMSEA 90% CI", "SRMR"),
  Value = c(
    round(fit_measures["chisq"], 2),
    fit_measures["df"],
    round(fit_measures["pvalue"], 4),
    round(fit_measures["cfi"], 3),
    round(fit_measures["tli"], 3),
    round(fit_measures["rmsea"], 3),
    paste0("[", round(fit_measures["rmsea.ci.lower"], 3), ", ",
           round(fit_measures["rmsea.ci.upper"], 3), "]"),
    round(fit_measures["srmr"], 3)
  ),
  Threshold = c("-", "-", "> 0.05", ">= 0.90", ">= 0.90", "<= 0.08", "-", "<= 0.08")
)

fit_df %>%
  kable(caption = "FAIR Model Fit Indices") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## Path Coefficients

```{r}
paths <- standardizedSolution(fit_fair) %>%
  filter(op == "~") %>%
  select(outcome = lhs, predictor = rhs, beta = est.std, se, z = z, pvalue) %>%
  mutate(
    outcome_label = construct_labels[outcome],
    predictor_label = construct_labels[predictor],
    significant = pvalue < 0.05,
    sig_marker = case_when(
      pvalue < 0.001 ~ "***",
      pvalue < 0.01 ~ "**",
      pvalue < 0.05 ~ "*",
      pvalue < 0.10 ~ ".",
      TRUE ~ ""
    )
  )

paths %>%
  mutate(
    beta = round(beta, 3),
    se = round(se, 3),
    z = round(z, 3),
    pvalue = round(pvalue, 4)
  ) %>%
  select(
    Outcome = outcome_label,
    Predictor = predictor_label,
    Beta = beta,
    SE = se,
    z = z,
    `p-value` = pvalue,
    Sig = sig_marker
  ) %>%
  kable(caption = "FAIR Model Path Coefficients (Standardized)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  footnote(general = "Significance: *** p < 0.001, ** p < 0.01, * p < 0.05, . p < 0.10")
```

## Path Coefficients Visualization

```{r fig.width=10, fig.height=6}
paths_plot <- paths %>%
  mutate(
    path = paste(predictor_label, "->", outcome_label),
    path = factor(path, levels = rev(path))
  )

ggplot(paths_plot, aes(x = path, y = beta)) +
  geom_col(aes(fill = significant), width = 0.7) +
  geom_errorbar(aes(ymin = beta - 1.96*se, ymax = beta + 1.96*se), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  coord_flip() +
  scale_fill_manual(
    values = c("TRUE" = "steelblue", "FALSE" = "gray70"),
    labels = c("TRUE" = "p < 0.05", "FALSE" = "Not significant"),
    name = NULL
  ) +
  labs(
    title = "FAIR Model: Standardized Path Coefficients",
    subtitle = "Error bars show 95% confidence intervals",
    x = NULL,
    y = "Standardized Beta"
  ) +
  theme(
    plot.title = element_text(face = "bold"),
    legend.position = "bottom"
  )
```

## Model Diagram

```{r fig.width=12, fig.height=10}
semPaths(
  fit_fair,
  whatLabels = "std",
  layout = "tree2",
  edge.label.cex = 0.8,
  node.label.cex = 0.7,
  sizeMan = 4,
  sizeLat = 8,
  style = "lisrel",
  nCharNodes = 0,
  nCharEdges = 0,
  fade = FALSE,
  residuals = FALSE,
  intercepts = FALSE,
  thresholds = FALSE,
  exoCov = FALSE
)
```

## R-squared Values

```{r}
r2 <- lavInspect(fit_fair, "rsquare")
r2_df <- data.frame(
  Construct = construct_labels[names(r2)],
  R_squared = round(r2, 3)
) %>%
  filter(!is.na(R_squared))

r2_df %>%
  arrange(desc(R_squared)) %>%
  kable(caption = "Variance Explained (R-squared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

# Modified FAIR Model (6 Constructs)

Combines Perceived Ease of Use and Perceived Usefulness into "Perceived Effectiveness".

```{r}
constructs_modified <- constructs
constructs_modified$perceived_effectiveness <- c(
  constructs$perceived_ease_of_use,
  constructs$perceived_usefulness
)
constructs_modified$perceived_ease_of_use <- NULL
constructs_modified$perceived_usefulness <- NULL

construct_labels_modified <- c(
  construct_labels,
  perceived_effectiveness = "Perceived Effectiveness"
)

modified_paths <- c(
  "# Fairness paths",
  "perceived_fairness ~ system_quality + service_quality",
  "",
  "# TAM paths (modified)",
  "perceived_effectiveness ~ system_quality + service_quality + perceived_fairness",
  "attitude_toward_use ~ perceived_fairness + perceived_effectiveness",
  "behavioral_intention ~ attitude_toward_use"
)

modified_model <- build_sem_model(constructs_modified, modified_paths)
cat("Modified FAIR Model Specification:\n\n")
cat(modified_model)
```

## Fit Modified Model

```{r}
fit_modified <- sem(
  modified_model,
  data = df_items,
  std.lv = TRUE,
  estimator = "MLM"
)
```

## Modified Model Fit

```{r}
if (lavInspect(fit_modified, "converged")) {
  fit_measures_mod <- fitMeasures(fit_modified, c(
    "chisq", "df", "pvalue",
    "cfi", "tli", "rmsea", "srmr"
  ))

  fit_df_mod <- data.frame(
    Index = c("Chi-square", "df", "p-value", "CFI", "TLI", "RMSEA", "SRMR"),
    Value = c(
      round(fit_measures_mod["chisq"], 2),
      fit_measures_mod["df"],
      round(fit_measures_mod["pvalue"], 4),
      round(fit_measures_mod["cfi"], 3),
      round(fit_measures_mod["tli"], 3),
      round(fit_measures_mod["rmsea"], 3),
      round(fit_measures_mod["srmr"], 3)
    ),
    Threshold = c("-", "-", "> 0.05", ">= 0.90", ">= 0.90", "<= 0.08", "<= 0.08")
  )

  fit_df_mod %>%
    kable(caption = "Modified FAIR Model Fit Indices") %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
} else {
  cat("WARNING: Modified FAIR model did not converge. Fit indices not available.\n")
  fit_measures_mod <- NULL
}
```

## Modified Model Path Coefficients

```{r}
if (lavInspect(fit_modified, "converged")) {
  paths_mod <- standardizedSolution(fit_modified) %>%
    filter(op == "~") %>%
    select(outcome = lhs, predictor = rhs, beta = est.std, se, pvalue) %>%
    mutate(
      outcome_label = construct_labels_modified[outcome],
      predictor_label = construct_labels_modified[predictor],
      significant = pvalue < 0.05,
      sig_marker = case_when(
        pvalue < 0.001 ~ "***",
        pvalue < 0.01 ~ "**",
        pvalue < 0.05 ~ "*",
        pvalue < 0.10 ~ ".",
        TRUE ~ ""
      )
    )

  paths_mod %>%
    mutate(
      beta = round(beta, 3),
      se = round(se, 3),
      pvalue = round(pvalue, 4)
    ) %>%
    select(
      Outcome = outcome_label,
      Predictor = predictor_label,
      Beta = beta,
      SE = se,
      `p-value` = pvalue,
      Sig = sig_marker
    ) %>%
    kable(caption = "Modified FAIR Model Path Coefficients (Standardized)") %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
} else {
  paths_mod <- NULL
  cat("Path coefficients not available (model did not converge).\n")
}
```

# Model Comparison

```{r}
if (!is.null(fit_measures_mod)) {
  comparison_df <- data.frame(
    Metric = c("Chi-square", "df", "CFI", "TLI", "RMSEA", "SRMR", "# Parameters"),
    FAIR = c(
      round(fit_measures["chisq"], 2),
      fit_measures["df"],
      round(fit_measures["cfi"], 3),
      round(fit_measures["tli"], 3),
      round(fit_measures["rmsea"], 3),
      round(fit_measures["srmr"], 3),
      fitmeasures(fit_fair, "npar")
    ),
    Modified_FAIR = c(
      round(fit_measures_mod["chisq"], 2),
      fit_measures_mod["df"],
      round(fit_measures_mod["cfi"], 3),
      round(fit_measures_mod["tli"], 3),
      round(fit_measures_mod["rmsea"], 3),
      round(fit_measures_mod["srmr"], 3),
      fitmeasures(fit_modified, "npar")
    )
  )

  comparison_df %>%
    kable(caption = "Model Comparison") %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
} else {
  cat("Model comparison not available (Modified FAIR model did not converge).\n")
}
```

# Hypothesis Testing Summary

```{r}
hypotheses <- data.frame(
  Hypothesis = paste0("H", 1:12),
  Path = c(
    "System Quality -> Perceived Fairness",
    "Service Quality -> Perceived Fairness",
    "Perceived Fairness -> Perceived Usefulness",
    "Perceived Fairness -> Attitude Toward Use",
    "System Quality -> Perceived Ease of Use",
    "Service Quality -> Perceived Ease of Use",
    "System Quality -> Perceived Usefulness",
    "Service Quality -> Perceived Usefulness",
    "Perceived Ease of Use -> Perceived Usefulness",
    "Perceived Ease of Use -> Attitude Toward Use",
    "Perceived Usefulness -> Attitude Toward Use",
    "Attitude Toward Use -> Behavioral Intention"
  )
)

# Match with results
hypotheses$Beta <- NA
hypotheses$p_value <- NA
hypotheses$Supported <- NA

for (i in 1:nrow(hypotheses)) {
  # Find matching path in results
  path_parts <- strsplit(hypotheses$Path[i], " -> ")[[1]]
  pred <- names(construct_labels)[construct_labels == path_parts[1]]
  out <- names(construct_labels)[construct_labels == path_parts[2]]

  match <- paths %>% filter(predictor == pred, outcome == out)

  if (nrow(match) == 1) {
    hypotheses$Beta[i] <- round(match$beta, 3)
    hypotheses$p_value[i] <- round(match$pvalue, 4)
    hypotheses$Supported[i] <- ifelse(match$pvalue < 0.05 & match$beta > 0, "Yes",
                                       ifelse(match$pvalue < 0.10 & match$beta > 0, "Marginal", "No"))
  }
}

hypotheses %>%
  kable(caption = "Hypothesis Testing Results (FAIR Model)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  row_spec(which(hypotheses$Supported == "Yes"), background = "#d4edda") %>%
  row_spec(which(hypotheses$Supported == "Marginal"), background = "#fff3cd") %>%
  row_spec(which(hypotheses$Supported == "No"), background = "#f8d7da")
```

# Conclusions

```{r results='asis'}
n_supported <- sum(hypotheses$Supported == "Yes", na.rm = TRUE)
n_marginal <- sum(hypotheses$Supported == "Marginal", na.rm = TRUE)
n_not_supported <- sum(hypotheses$Supported == "No", na.rm = TRUE)

cat("## Summary\n\n")
cat("- **Supported (p < 0.05):**", n_supported, "hypotheses\n")
cat("- **Marginally supported (p < 0.10):**", n_marginal, "hypotheses\n")
cat("- **Not supported:**", n_not_supported, "hypotheses\n\n")

cat("## Key Findings\n\n")

# Report significant paths
sig_paths <- paths %>% filter(pvalue < 0.05)
if (nrow(sig_paths) > 0) {
  cat("**Significant paths (p < 0.05):**\n\n")
  for (i in 1:nrow(sig_paths)) {
    cat("- ", sig_paths$predictor_label[i], " -> ", sig_paths$outcome_label[i],
        " (beta = ", round(sig_paths$beta[i], 3), ", p = ", round(sig_paths$pvalue[i], 4), ")\n", sep = "")
  }
}

# Report non-significant paths
nonsig_paths <- paths %>% filter(pvalue >= 0.05)
if (nrow(nonsig_paths) > 0) {
  cat("\n**Non-significant paths (p >= 0.05):**\n\n")
  for (i in 1:nrow(nonsig_paths)) {
    cat("- ", nonsig_paths$predictor_label[i], " -> ", nonsig_paths$outcome_label[i],
        " (beta = ", round(nonsig_paths$beta[i], 3), ", p = ", round(nonsig_paths$pvalue[i], 4), ")\n", sep = "")
  }
}

cat("\n## Limitations\n\n")
cat("- Item selection was based on psychometric criteria (item-total correlation, factor loadings)\n")
cat("- No split-sample validation was performed\n")
cat("- Model fit indices should be evaluated against recommended thresholds\n")
```

# Save Results

```{r}
saveRDS(fit_fair, "fit_fair.rds")
saveRDS(fit_modified, "fit_modified.rds")
saveRDS(paths, "paths_fair.rds")
saveRDS(paths_mod, "paths_modified.rds")
cat("Results saved.\n")
```
