---
title: "Step 4: Post-Hoc Modification Indices"
output:
  html_notebook:
    toc: true
    toc_float: true
---

This analysis extracts modification indices from the FAIR SEM model to identify
sources of misfit and tests theoretically motivated alternative specifications.

```{r setup, message=FALSE, warning=FALSE}
library(lavaan)
library(semTools)
library(dplyr)
library(knitr)
library(kableExtra)
```

# Load Data

```{r}
constructs <- readRDS("constructs_final.rds")
df_items <- readRDS("df_items.rds")

construct_labels <- c(
  system_quality = "System Quality",
  service_quality = "Service Quality",
  perceived_fairness = "Perceived Fairness",
  perceived_ease_of_use = "Perceived Ease of Use",
  perceived_usefulness = "Perceived Usefulness",
  attitude_toward_use = "Attitude Toward Use",
  behavioral_intention = "Behavioral Intention"
)

cat("Sample size:", nrow(df_items), "\n")
```

# Rebuild and Fit FAIR Model

```{r}
build_sem_model <- function(constructs, structural_paths) {
  measurement <- sapply(names(constructs), function(name) {
    items <- constructs[[name]]
    paste0(name, " =~ ", paste(items, collapse = " + "))
  })
  paste(c(measurement, "", "# Structural paths", structural_paths), collapse = "\n")
}

fair_paths <- c(
  "perceived_fairness ~ system_quality + service_quality",
  "perceived_ease_of_use ~ system_quality + service_quality",
  "perceived_usefulness ~ system_quality + service_quality + perceived_fairness + perceived_ease_of_use",
  "attitude_toward_use ~ perceived_fairness + perceived_ease_of_use + perceived_usefulness",
  "behavioral_intention ~ attitude_toward_use"
)

fair_model <- build_sem_model(constructs, fair_paths)
fit_fair <- sem(fair_model, data = df_items, std.lv = TRUE, estimator = "MLM")
```

## Current Model Fit

```{r}
fit_measures <- fitMeasures(fit_fair, c("cfi", "tli", "rmsea", "srmr"))

data.frame(
  Index = c("CFI", "TLI", "RMSEA", "SRMR"),
  Value = round(as.numeric(fit_measures), 3),
  Threshold = c(">= 0.90", ">= 0.90", "<= 0.08", "<= 0.08")
) %>%
  kable(caption = "Current FAIR Model Fit") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

# Modification Indices

```{r}
mi <- modificationIndices(fit_fair, sort. = TRUE, minimum.value = 3.84)

head(mi, 30) %>%
  select(lhs, op, rhs, mi, epc, sepc.all) %>%
  mutate(mi = round(mi, 1), epc = round(epc, 3), sepc.all = round(sepc.all, 3)) %>%
  kable(caption = "Top 30 Modification Indices") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## Cross-Loadings

Items that want to load on constructs other than their assigned one.

```{r}
mi_crossload <- mi %>%
  filter(op == "=~") %>%
  head(15) %>%
  mutate(construct = construct_labels[lhs])

mi_crossload %>%
  select(Construct = construct, Item = rhs, MI = mi, EPC = epc, `Std EPC` = sepc.all) %>%
  mutate(MI = round(MI, 1), EPC = round(EPC, 3), `Std EPC` = round(`Std EPC`, 3)) %>%
  kable(caption = "Top Cross-Loading Suggestions") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## Error Covariances

Pairs of items whose residuals are correlated beyond what the model explains.

```{r}
mi_errcov <- mi %>%
  filter(op == "~~", !lhs %in% names(construct_labels), !rhs %in% names(construct_labels)) %>%
  head(15)

reverse_items <- c("Q2", "Q7", "Q9", "Q15", "Q25", "Q31", "Q37", "Q42")

mi_errcov %>%
  mutate(
    Both_Reversed = (lhs %in% reverse_items) & (rhs %in% reverse_items)
  ) %>%
  select(Item1 = lhs, Item2 = rhs, MI = mi, `Both Reversed` = Both_Reversed) %>%
  mutate(MI = round(MI, 1)) %>%
  kable(caption = "Top Error Covariance Suggestions") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

# Method Factor Model

Adding an orthogonal method factor for the reverse-coded items still in the model
to capture shared response bias from negative item wording.

```{r}
paths_orig <- standardizedSolution(fit_fair) %>%
  filter(op == "~") %>%
  select(lhs, rhs, beta_orig = est.std, p_orig = pvalue)

items_in_model <- unlist(constructs)
reverse_in_model <- intersect(reverse_items, items_in_model)
cat("Reverse-coded items in model:", paste(reverse_in_model, collapse = ", "), "\n")

method_factor <- paste0("method =~ ", paste(reverse_in_model, collapse = " + "))

fair_model_method <- paste(c(fair_model, "",
  "# Method factor for reverse-coded items",
  method_factor,
  "method ~~ 0*system_quality",
  "method ~~ 0*service_quality",
  "method ~~ 0*perceived_fairness",
  "method ~~ 0*perceived_ease_of_use",
  "method ~~ 0*perceived_usefulness",
  "method ~~ 0*attitude_toward_use",
  "method ~~ 0*behavioral_intention"
), collapse = "\n")

fit_method <- suppressWarnings(
  sem(fair_model_method, data = df_items, std.lv = TRUE, estimator = "MLM")
)
```

## Method Factor Model Fit

```{r}
fit_method_measures <- fitMeasures(fit_method, c("cfi", "tli", "rmsea", "srmr"))

data.frame(
  Index = c("CFI", "TLI", "RMSEA", "SRMR"),
  Original = round(as.numeric(fit_measures), 3),
  Method_Factor = round(as.numeric(fit_method_measures), 3),
  Change = round(as.numeric(fit_method_measures) - as.numeric(fit_measures), 3)
) %>%
  kable(caption = "Fit Comparison: Original vs Method Factor Model") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## Path Coefficients Comparison

```{r}
paths_method <- standardizedSolution(fit_method) %>%
  filter(op == "~") %>%
  select(lhs, rhs, beta_method = est.std, p_method = pvalue)

comparison <- merge(paths_orig, paths_method, by = c("lhs", "rhs"))
comparison$lhs_label <- construct_labels[comparison$lhs]
comparison$rhs_label <- construct_labels[comparison$rhs]

comparison %>%
  select(
    `Path To` = lhs_label, `Path From` = rhs_label,
    `Beta (Orig)` = beta_orig, `Beta (Method)` = beta_method,
    `p (Orig)` = p_orig, `p (Method)` = p_method
  ) %>%
  mutate(across(where(is.numeric), ~round(., 3))) %>%
  kable(caption = "Path Coefficients: Original vs Method Factor Model") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## Method Factor Loadings

```{r}
standardizedSolution(fit_method) %>%
  filter(op == "=~", lhs == "method") %>%
  select(Item = rhs, Loading = est.std, `p-value` = pvalue) %>%
  mutate(Loading = round(Loading, 3), `p-value` = round(`p-value`, 4)) %>%
  kable(caption = "Method Factor Loadings") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

# Direct PU -> BI Path

The original TAM (Davis, 1989) includes a direct path from Perceived Usefulness
to Behavioral Intention, which the FAIR model omits.

```{r}
fair_paths_pu_bi <- c(
  "perceived_fairness ~ system_quality + service_quality",
  "perceived_ease_of_use ~ system_quality + service_quality",
  "perceived_usefulness ~ system_quality + service_quality + perceived_fairness + perceived_ease_of_use",
  "attitude_toward_use ~ perceived_fairness + perceived_ease_of_use + perceived_usefulness",
  "behavioral_intention ~ attitude_toward_use + perceived_usefulness"
)

fair_model_pu_bi <- build_sem_model(constructs, fair_paths_pu_bi)

fit_pu_bi <- suppressWarnings(
  sem(fair_model_pu_bi, data = df_items, std.lv = TRUE, estimator = "MLM")
)

if (lavInspect(fit_pu_bi, "converged")) {
  fit_pu_bi_measures <- fitMeasures(fit_pu_bi, c("cfi", "tli", "rmsea", "srmr"))

  data.frame(
    Index = c("CFI", "TLI", "RMSEA", "SRMR"),
    Original = round(as.numeric(fit_measures), 3),
    PU_BI = round(as.numeric(fit_pu_bi_measures), 3),
    Change = round(as.numeric(fit_pu_bi_measures) - as.numeric(fit_measures), 3)
  ) %>%
    kable(caption = "Fit Comparison: Original vs PU -> BI Model") %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
} else {
  cat("Model did not converge.\n")
}
```

```{r results='asis'}
# Check for identification issues
vcov_ok <- tryCatch({
  vcov(fit_pu_bi)
  TRUE
}, error = function(e) FALSE)

if (!vcov_ok) {
  cat("**Note:** The information matrix could not be inverted, indicating model",
      "identification problems. Standard errors are not computable. This is likely",
      "because Attitude Toward Use already predicts Behavioral Intention at the",
      "boundary (beta = 1.000), leaving no identifiable variance for a second",
      "predictor. The original FAIR model specification is retained.\n")
}
```
